---
title: 'Assignment 2'
author: "Madison Wozniak"
date: "9/9/21"
output:
  pdf_document: default
  html_document: default
  word_document: default
---


<center> <h1>

Assignment 2</h1> </center>

You are asked to submit both the R Markdown file and its pdf output.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<span style="color:red">**Q1.** </span> Using R, compute the following
$$\frac{0.35-0.3}{\sqrt{\frac{0.2*(1-0.4)}{50}}}$$ 

<span style="color:red">**Answer** </span>

```{r}
(0.35-0.3)/(sqrt((0.2*(1-0.4))/(50)))
```

<span style="color:red">**Q2.** </span> Define two variables`weight = [150, 160, 170, 120]` and `height = [5.2, 5.5, 5.8, 5.2]` and plot weight vs height.

```{r}
weight<-c(150,160,170,120)
height<-c(5.2,5.5,5.8,5.2)
plot(weight,height)
```


<span style="color:red">**Q3.** </span> Without running any code, predict the outcome of each line:

x <- c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29) 

x[1:5]

x<-c(1,2,3,4,5)

x[c(1, 4)]

Select rows 1 and 4

x[-c(2, 5)]

Remove rows 2 and 5

 <pre><code><span style="color:blue"> x <- c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29)  </span>

 <span style="color:blue"> x[1:5]</span>

 <span style="color:blue">  x[c(1, 4)]</span>

 <span style="color:blue">  x[-c(2, 5)]</span> </code></pre>
 


<span style="color:red">**Q4.** </span>Generate the matrix below

\[
\begin{bmatrix}
   1       & 4 & 7 & 3 \\
    2      & 4 & 3 &  8 \\
    3      & 2 & 1 & 5
\end{bmatrix}
\]
```{r}
G<-matrix(c(1,2,3,4,4,2,7,3,1,3,8,5),3,4)
G
```

(a) Report the 2nd and the 3rd row.
```{r}
G[c(2,3),]
```

(b) Report all columns except the 2nd one.
```{r}
G[,-2]
```

(c) Rename row and column names to your names of choice. 
```{r}
rownames(G)<-c("lions","tigers","bears")
colnames(G)<-c("pink","purple","green","gold")
```

(d) Call the second and the third row using the names you defined.
```{r}
G[c("tigers","bears"),]
```

<span style="color:red">**Q5.** </span> Create a dataframe with four features (columns), first is called `no_bedrooms`, `location`, `age`, `price`. Here is the info for five houses:

House 1: 4, 'Boston', 35, $500K

House 2: 1, 'San Francisco', 55, $900K

House 3: 4, 'Hartford', 87, $300K

House 4: 3, 'Houston', 45, $280K

House 5: 3, 'Seattle', 35, $850K

```{r}
no_bedrooms<-c(4,1,4,3,3)
location<-c("Boston","San Francisco","Hartford","Houston","Seattle")
age<-c(35,55,87,45,35)
price<-c(500000,900000,300000,280000,850000)
df<-data.frame(no_bedrooms,location,age,price)
df
```

<span style="color:red">**Q6.** </span>Load the dataset that is already built-in data in R: `data(mtcars)`.

(a) How many observations are there in this dataset?
```{r}
data(mtcars)
mtcars
str(mtcars)
```
32 observations
(b) How many variables does this dataset have? What are the names of these variables?

11 variables: `mpg,cyl,disp,hp,drat,wt,qsec,vs,am,gear,carb`

(c) What are the classes of the variables?
```{r}
class(mtcars$mpg)
class(mtcars$cyl)
class(mtcars$disp)
class(mtcars$hp)
class(mtcars$drat)
class(mtcars$wt)
class(mtcars$qsec)
```

(d) What are the statistics of `mpg` such as mean, median, standard deviation.
```{r}
mean(mtcars$mpg)
median(mtcars$mpg)
sd(mtcars$mpg)
```

(e) How many observations are with `drat`> 3?
```{r}
log_drat<-(mtcars$drat>3)
```

<span style="color:red">**Q7.** </span> Start by loading the library and data:
<pre><code>
library(dslabs)
data(murders) 
</code></pre>
```{r}
library(dslabs)
data(murders)
murders
```

(a) Compute the per 100,000 murder rate for each state and store it in an object called `murder_rate`. Then use logical operators to create a logical vector named low that tells us which entries of murder_rate are lower than 1.
```{r}
murder_rate<-(murders$total)/(murders$population)*100000
murder_rate
low<-murder_rate<1
low
```

(b) Now use the results from the previous exercise and the function which to determine the indices of murder_rate associated with values lower than 1.
```{r}
which(low)
```

(c) Use the results from the previous exercise to report the names of the states with murder rates lower than 1.
```{r}
murders$state[low]
```

(d) Now extend the code from exercises 2 and 3 to report the states in the Northeast with murder rates lower than 1. Hint: use the previously defined logical vector low and the logical operator &.
```{r}
murders$state[low&(murders$region=="Northeast")]
```

(e) In a previous exercise we computed the murder rate for each state and the average of these numbers. How many states are below the average?
```{r}
mean(murder_rate)
murders$state[mean(murder_rate)]
```

(f) Use the match function to identify the states with abbreviations AK, MI, and IA. Hint: start by defining an index of the entries of murders$abb that match the three abbreviations, then use the [ operator to extract the states.
```{r}
murders$abb
match(c("AK","MI","IA"),murders$abb)
```

(g) Use the %in% operator to create a logical vector that answers the question: which of the following are actual abbreviations: MA, ME, MI, MO, MU?
```{r}
murders$abb%in%c("MA","ME","MI","MO","MU")
```

(h) Extend the code you used in exercise 7 to report the one entry that is not an actual abbreviation. Hint: use the ! operator, which turns FALSE into TRUE and vice versa, then which to obtain an index.
```{r}
!murders$abb%in%c("MA","ME","MI","MO","MU")
```

(i)  Create a histogram of the state populations.
```{r}
hist(murders$population)
```

(j) Generate boxplots of the state populations by region.
```{r}
boxplot(population~region,data=murders)
```

