---
title: 'Madison Wozniak: Assignment 3'
output:
  pdf_document: default
  html_document: default
  word_document: default
---


# Assignment 3

You are asked to submit both the R Markdown file and its pdf output.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<span style="color:red">**Q1.** </span> Write an if-else statement:

1. If the number is greater than 0 and less than 10, print: "This number is between 0 and 10"

2. If the number is greater than 10, print: "This number is greater than 10"

3. If the number is less than 0, print: "This number is a negative number"

4. Otherwise print: "This number is either 0 or 10"
```{r}
x<-12
if(x>0){
  print("This number is between 0 and 10")
}else if(x<10){
  print("This number is between 0 and 10")
} else if(x>10){
  print("This number is greater than 10")
}else if(x<0){
  print("This number is a negative number")
}else{
  print("This number is either 0 or 10")
} 

```


<span style="color:red">**Q2.** </span> Write a function that gets a vector as its input and returns the mean and the standard deviation of the vector using the formula below:

$$\left(\frac{\sum (x_i-\bar{x})^2}{n-1} \right)^{1/2}$$ 
where $\bar{x}$ is the mean and n is the length of the vector.
```{r}
x<-c(1,5,4,9,1)
 n<-length(x)
((sum(x-mean(x))^2)/(n-1))^(1/2)
sd_by_hand<-function(x){
  
  x_bar<-mean(x)
  n<-length(x)
  sd_<-(((sum(x-x_bar^2))/(n-1))^(1/2))
  mean_<-mean(x)
  c(mean_,sd_)
    
}
sd_by_hand(c(1,5,4,9,1))
```

<span style="color:red">**Q3.** </span> 

(a) Write a function that returns Euclidian Distance between two k-dimensional vectors:

$$d=\sqrt{(x_1-y_1)^2+(x_2-y_2)^2+...+(x_k-y_k)^2}$$
```{r}
Euclidian_Distance<-function(x){
  sqrt((sum(x-y))^2)
}
x<-c(4,5,7,3,56)
y<-c(3,4,5,6,7)
Euclidian_Distance(x)
```

(b) Write a function that will input the vectors x, y and p and will return the distance between two k-dimensional vectors:

$$d_p(x,y)=((x_1-y_1)^p+(x_2-y_2)^p+...+(x_k-y_k)^p)^{1/p}$$

Pick the default value for p as 2.
```{r}
distance_function<-function(x,y,p){
  p<-2
  x<-c(3,4,5,6,7)
y<-c(1,2,3,4,5)
x_k<-length(x)
y_k<-length(y)
(sum((x-y)^p))^(1/p)
}
distance_function(x,y,p)
```

<span style="color:red">**Q4.** </span>  Create a function altman_plot that takes two arguments, `x` and `y`, and plots the difference against the sum.
```{r}
altman_plot<-function(x,y){
  plot(x+y,x-y)
}
x<-c(1,2,4,7,8)
y<-c(3,5,2,9,1)
altman_plot(x,y)
```

<span style="color:red">**Q5.** </span>  Write a function `compute_s_n_2` that for any 
given n, computes the sum:

$$1^2+2^2+3^2+4^2+...+(n-1)^2+n^2$$
```{r}
compute_s_n_2<-function(n){
  x<-1:n
  sum(x)
}
```

(a) Find `compute_s_n_2(30)`.
```{r}
compute_s_n_2<-function(n){
  x<-1:n
  sum(x)
}
compute_s_n_2(30)
```

(b) Confirm that the formula for this sum is $=\frac{n(n+1)(2n+1)}{6}$.


<span style="color:red">**Q6.** </span> Which of the following built-in datasets is tidy (you can pick more than one):

(a) BJsales
(b) EuStockMarkets
(c) DNase
(d) Formaldehyde
(e) Orange
(f) UCBAdmissions

```{r}
head(BJsales)
# not tidy
head(EuStockMarkets)
#not tidy
head(DNase)
#tidy
head(Formaldehyde)
#tidy
head(Orange)
#tidy
head(UCBAdmissions)
#tidy
```


<span style="color:red">**Q7.** </span> Load the `dplyr` package and the `murders` dataset.
```{r}
library(tidyverse)
library(dslabs)
murders
```
(a) By using dplyr's `mutate` function, add a new column:

`population_in_millions = population / 10^6`
```{r}
murders_added<-mutate(murders,population_in_millions=population/(10^6))
murders_added
```


(b) If `rank(x)` gives you the ranks of `x` from lowest to highest, `rank(-x)` gives you the ranks from highest to lowest. Use the function mutate to add a column rank containing the rank, from highest to lowest murder rate. Make sure you redefine murders so we can keep using this variable.
```{r}
murders_ranked<-murders%>%mutate(murd_rank=rank(-(total/(population)*100000)))
murders_ranked
```

(c) Select the columns `state`, `population` and give it a name `new_df`.
```{r}
library(tidyverse)
new_df<-murders%>%
  select(state,population)
new_df
```

(d) Filter the observations with only `state== 'New York'` and call it as `murders_ny`.
```{r}
murders_ny<-murders%>%
  filter(state=="New York")
murders_ny
```

(e) Remove all the observations with `state=='Florida'` and call that dataframe as `murders_no_fl`.
```{r}
murders_no_fl<-murders%>%
filter(state!="Florida")
murders_no_fl
```

(f) Filter the murders dataset using `%in%` to filter the observations with `state=='New York'` or `state=='Texas'`.
```{r}
murders_tex_ny<-murders%>%
  filter(state%in%c("New York"))
murders_tex_ny
```

(g) Suppose you want to live in the Northeast or West and want the murder rate to be less than 1. How many options do you have?
```{r}
safe_northeast_west<-murders%>%
  mutate(rate=(total/population*100000))%>%
  filter((region%in%c("Northeast","West")&(rate<1)))
safe_northeast_west
```

<span style="color:red">**Q8.** </span> Use a pipe to create a new data frame called my_states that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns.

```{r}
my_states <- murders %>%
 mutate(rate=(total/population*100000),murd_rank=rank(-(total/population*100000))) %>%
filter((region%in%c("Northeast","West")&(rate<1))) %>%
select(state,rate,murd_rank)
my_states
```

<span style="color:red">**Q9.** </span> Install the NHANES package, load the data `NHANES`. 
```{r}
#install.packages("NHANES")
library(NHANES)
data(NHANES)
```

Observe that `NHANES` data has many missing values. 

(a) Find the mean and the standard deviation of the variable `Age`. Remember the exclude the missing values. Hint: Add `na.rm = TRUE` inside the mean and the sd functions.
```{r}
mean(NHANES$Age,na.rm=TRUE)
sd(NHANES$Age,na.rm=TRUE)
```

(b) First select the group as 20-to-29-year-old females. `AgeDecade` is a categorical variable with these ages. Note that the category is coded like " 20-29", with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the `BPSysAve` variable? Save it to a variable called `ref`.
```{r}
filtered_age<-NHANES%>%
  filter(AgeDecade==" 20-29")
filtered_age
ref<-filtered_age%>%
  filter(BPSysAve!="NA")%>%
  summarize(avg=mean(BPSysAve,na.rm=TRUE),standev=sd(BPSysAve,na.rm=TRUE))
ref
```

*Hint:* Use `filter` and `summarize` and use the `na.rm = TRUE` argument when computing the average and standard deviation. You can also `filter` the NA values using `filter`.

(c) Using a pipe, assign the average to a numeric variable `ref_avg`. Hint: Use the code similar to above and then `pull`.
```{r}
ref_avg<-filtered_age%>%
  filter(BPSysAve!="NA")%>%
  summarize(avg=mean(BPSysAve,na.rm=TRUE),standev=sd(BPSysAve,na.rm=TRUE))%>%
  pull(avg)
ref_avg
```

(d) Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in the earlier question. Note that the age groups are defined by AgeDecade. Hint: rather than filtering by age and gender, `filter` by *Gender* and then use `group_by`.
```{r}
filtered_female_age<-NHANES%>%
  filter(Gender=="female")%>%
  group_by(AgeDecade)%>%
  filter(BPSysAve!="NA")%>%
  summarize(avg=mean(BPSysAve,na.rm=TRUE),standev=sd(BPSysAve,na.rm=TRUE))
filtered_female_age
```

(e) Repeat exercise (d) for males.
```{r}
filtered_male_age<-NHANES%>%
  filter(Gender=="male")%>%
  group_by(AgeDecade)%>%
  filter(BPSysAve!="NA")%>%
  summarize(avg=mean(BPSysAve,na.rm=TRUE),standev=sd(BPSysAve,na.rm=TRUE))
filtered_male_age
```

(f) We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because `group_by` permits us to group by more than one variable. Obtain one big summary table using `group_by(AgeDecade, Gender)`.
```{r}
combined_summary<-NHANES%>%
group_by(AgeDecade,Gender)%>%
  summarize(avg=mean(BPSysAve,na.rm=TRUE),standev=sd(BPSysAve,na.rm=TRUE))
combined_summary
```

(g)  For males between the ages of 40-49, compare systolic blood pressure `BPSysAve` across race as reported in the Race1 variable. Order the resulting table from lowest to highest average systolic blood pressure.
```{r}
male_filter_age<-NHANES%>%
  filter(AgeDecade==" 40-49")
male_comparison<-male_filter_age%>%
  summarize(Race1,BPSysAve)%>%
  arrange(BPSysAve)
male_comparison
```


