---
title: "CaseStudy2"
author: "Phia Rau Halleen, Cassia,  Madison Wozniak"
date: "November 2, 2021"
output:
  pdf_document: default
  html_document: default
---

<center> <h1>

BSDS 100-Case Study 2</h1> </center>

You are asked to submit both the R Markdown file and its pdf output.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<span style="color:red">**Q1.** </span> Install the packages `nycflights13`,  `maps`` and load the datasets: 

```{r}
library(tidyverse)
library(maps)
library(nycflights13)
data(flights)
data(airlines)
data(airports)
data(planes)
data(weather)
```
Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

a) Use the size or color of the points to display the average delay for each airport.
```{r}
library(dplyr)
avg_delay <- flights%>%
  group_by(dest)%>%
  summarise(delay = mean(arr_delay, na.rm = TRUE))%>%
  inner_join(airports, by = c(dest = "faa"))

avg_delay %>%
  ggplot(aes(x = lon, y = lat, colour = delay)) + borders("state")  +   geom_point() + coord_quickmap() + scale_color_viridis_b()
```


b) Add the location of the origin and destination (i.e. the lat and lon) to flights.
```{r}
location <- airports%>%
  select(faa,lat,lon)

flights%>%
  select(year: day, hour, origin, dest)%>%
  left_join(location, by = c("origin" = "faa"))%>%
  left_join(location, by = c("dest" = "faa"))
```


c) Is there a relationship between the age of a plane and its delays?
```{r}
plane_age_delay <- inner_join(flights, select(planes, tailnum, plane_year = year), by = "tailnum") %>%
  mutate(age = year - plane_year) %>%
  filter(!is.na(age)) %>%
  mutate(age = if_else(age > 25, 25L, age)) %>%
  group_by(age) %>%
  summarise(dep_delay_mean = mean(dep_delay, na.rm = TRUE), 
            dep_delay_sd = sd(dep_delay, na.rm = TRUE), 
            arr_delay_mean = mean(arr_delay, na.rm = TRUE), 
            arr_delay_sd = sd(arr_delay, na.rm = TRUE),  
            n_arr_delay = sum(!is.na(arr_delay)), 
            n_dep_delay = sum(!is.na(dep_delay)))

ggplot(plane_age_delay, aes(x = age, y = dep_delay_mean)) +
  geom_point() +
  scale_x_continuous("Age of plane (years)", breaks = seq(0, 30, by = 10)) +
  scale_y_continuous("Mean Departure Delay (minutes)")
```
There is a relationship between the age of a plane and its delays.


d) What weather conditions make it more likely to see a delay?
```{r}
flight_weather <- flights %>% 
  inner_join(weather, by = c("origin" = "origin", "year" = "year", 
                             "month" = "month", "day" = "day", 
                             "hour" = "hour"))
flight_weather%>%
  group_by(precip)%>%
  summarise(delay = mean(dep_delay, na.rm = TRUE))%>%
  ggplot(aes(x = precip, y = delay)) +
  geom_line() + geom_point()

flight_weather %>%
  ungroup() %>%
  mutate(visib_cat = cut_interval(visib, n = 10)) %>%
  group_by(visib_cat) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = visib_cat, y = dep_delay)) + 
  geom_point()
```


e) What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.
```{r}
nycflights13::flights
June13_2013 <- flights%>%
  filter(year == 2013, month == 6, day == 13)%>%
  group_by(dest)%>%
  summarize(dep_delay = mean(arr_delay, na.rm = TRUE))%>%
  inner_join(airports, by = c("dest" = "faa"))%>%
ggplot(aes(x=lon,y=lat,))

June13_2013b <- flights%>%
  filter(year == 2013, month == 6, day == 13)%>%
  group_by(dest)%>%
  summarize(dep_delay = mean(arr_delay, na.rm = TRUE))%>%
  inner_join(airports, by = c("dest" = "faa"))%>%
  ggplot(aes(x=lon, y=lat, size=dep_delay, color=dep_delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() +
  scale_color_viridis_b()

June13_2013b
```




<span style="color:red">**Q2.** </span>  Using the `tweets.csv` data that is available on Canvas

(a) Identify all tweets with the word ‘flight’ in them.
```{r}
library(readxl)
library(htmlwidgets)
flight_tweets <- read.csv("tweets-1.csv", header=FALSE)
head(flight_tweets)
flight_tweets%>%
  filter(str_detect(V1,"flight")) 
```
There are three tweets with a question mark. 
(b) How many tweets end in a question mark?
```{r}

flight_tweets%>%
  filter(str_detect(V1,"\\?$"))

#the question mark is taken as a special expression, so we used the backslash, 
#then we use the "$" to specify the end of the string. 

```
There are two tweets with a question mark. 

(c) How many tweets have airport codes in them (assume any three subsequent capital letters are airport codes).
```{r}
str_match_all(tweets_1," [A-Z]{3} ")
```
There are 6 tweets that contain airport codes.

(d) Identify all tweets with URLs in them.
```{r}
view_tweets<-str_extract_all(tweets_1,"http[s]?://(?:[a-zA-Z]|[0-9]|[$-_@.&+]|[!*\\(\\),]|(?:%[0-9a-fA-F][0-9a-fA-F]))+")
view_tweets
```
There are 3 URLS across 2 tweets.

(e) Replace all instances of repeated exclamation points with a single exclamation point.
```{r}
replace_exclamations<-tweets_1%>%
mutate(str_replace_all(tweets_1,"\\!+","\\!"))
replace_exclamations
```

(f) Replace consecutive exclamation points, question marks, and periods with a single period, split the tweet on periods.
```{r}
replace_with_period<-tweets_1%>%
mutate(period_line_break=(str_replace_all(tweets_1,"[!+|?+|.+]",".")))
line_break<-replace_with_period%>%
  mutate(str_split(tweets_1,". "))
line_break
```
