---
title: "Case Study 2"
author: "Madison Wozniak"
date: "OCtober 31st, 2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---


<center> <h1>

BSDS 100-Case Study 2</h1> </center>

You are asked to submit both the R Markdown file and its pdf output.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<span style="color:red">**Q1.** </span> Install the packages `nycflights13`,  `maps`` and load the datasets: 

```{r}
library(tidyverse)
library(maps)
install.packages("maps")
library(nycflights13)
data(flights)
data(airlines)
data(airports)
data(planes)
data(weather)
```
Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

a) Use the size or color of the points to display the average delay for each airport.
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()

```

b) Add the location of the origin and destination (i.e. the lat and lon) to flights.

c) Is there a relationship between the age of a plane and its delays?

d) What weather conditions make it more likely to see a delay?

e) What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.


<span style="color:red">**Q2.** </span>  Using the `tweets.csv` data that is available on Canvas

(a) Identify all tweets with the word ‘flight’ in them.

(b) How many tweets end in a question mark?

(c) How many tweets have airport codes in them (assume any three subsequent capital letters are airport codes).
```{r}
install.packages("tweets.csv")
str_match_all(tweets_1," [A-Z]{3} ")
```


(d) Identify all tweets with URLs in them.
```{r}
view_tweets<-str_extract_all(tweets_1,"http[s]?://(?:[a-zA-Z]|[0-9]|[$-_@.&+]|[!*\\(\\),]|(?:%[0-9a-fA-F][0-9a-fA-F]))+")
view_tweets
```

(e) Replace all instances of repeated exclamation points with a single exclamation point.
```{r}
replace_exclamations<-tweets_1%>%
mutate(str_replace_all(tweets_1,"\\!+","\\!"))
replace_exclamations
```

(f) Replace consecutive exclamation points, question marks, and periods with a single period, split the tweet on periods.
```{r}
replace_with_period<-tweets_1%>%
mutate(period_line_break=(str_replace_all(tweets_1,"[!+|?+|.+]",".")))
replace_with_period
line_break<-replace_with_period%>%
  mutate(str_split(tweets_1,". "))
line_break
```



